\appendix

% PROMPTS
\section{Prompts}
\label{app:prompts}
This section contains all the prompts used throughout this work to guide the behavior of LLM agents, including those for initialization, interaction, content generation, and opinion update.

\subsection{Agent roleplay}
\label{app:agent}
Before performing any action, agents are initialized with a detailed profile that defines their identity, including political orientation and current opinions, and provides them complete descriptions of the topics and the opinions held by their supported coalition.

\input{prompts/agent_roleplay}

% Base prompts
\subsection{Actions}
\label{app:prompt_actions}
The following are the prompts for the actions that agents can perform when they are active.
Please note that the prompts for \textit{post} and \textit{comment} refer to base agents, while those for misinformation agents are provided in the next subsection.

\subsubsection{Post}
\input{prompts/post}

\subsubsection{Comment}
\input{prompts/comment}

\subsubsection{Reaction}
\input{prompts/reaction}

\subsubsection{Evaluate follow}
\input{prompts/evaluate_follow}

% Misinfo
\subsection{Misinformation actions}
\label{app:prompt_misinfo}
Misinformation agents share most of the prompts with base agents, except for those related to content generation: \textit{post} and \textit{comment}.
In these cases, the 280 characters limit was enforced, since misinformation agents tended to include details and exceed the intended size.
They are encouraged to produce misleading content to support their view, even by inventing realistic data and citing studies.

\subsubsection{Post}
\input{prompts/misinfo_post}

\subsubsection{Comment}
\input{prompts/misinfo_comment}

% Opinion update
\subsection{Opinion update}
\label{app:prompt_opinion}
The prompt to update the opinion has two main purposes: updating the textual opinion, and assigning a stance label, later mapped to a numerical score.

It includes the topics to update, a bias instruction (stronger for misinformation agents), and a memory of the daily interactions to support context-aware updates,

They are also provided formatting guidelines to reduce errors and simplify the output extraction.

\input{prompts/opinion_update}


% COALITIONS
\section{Coalition opinions}
\label{app:coalition_opinions}
The following are the opinions of the coalitions considered in this work.
They also serve as the initial opinions for the supporting agents.

\subsection{Centre-Left}
\input{Coalitions/Centre-Left_opinions}

\subsection{Movimento 5 Stelle (M5S)}
\input{Coalitions/M5S_opinions}

\subsection{Right}
\input{Coalitions/Right_opinions}

\subsection{Third Pole}
\input{Coalitions/Third_Pole_opinions}
